---
title: "Titanic Challenge Kaggle"
bibliography: references.bib
---

```{r}
#| label: load-packages
#| include: false

library(tidyverse)
library(knitr)

library(randomForest)
library(nnet)
```

## The Challenge

The sinking of the Titanic is one of the most infamous shipwrecks in history.

On April 15, 1912, during her maiden voyage, the widely considered "unsinkable" RMS Titanic sank after colliding with an iceberg. Unfortunately, there weren't enough lifeboats for everyone onboard, resulting in the death of 1502 out of 2224 passengers and crew.

While there was some element of luck involved in surviving, it seems some groups of people were more likely to survive than others.

In this challenge, we ask you to build a predictive model that answers the question: "what sorts of people were more likely to survive?" using passenger data (ie name, age, gender, socio-economic class, etc). [@titanic]

## Loading Data

```{r}
#| output: false

train <- read_csv("./data/train.csv")
test  <- read_csv("./data/test.csv")
```

```{r}
#| echo: false

kable(head(train))
```

### Prepping Training Data

```{r}

titanic.training <- train %>% select(PassengerId, Survived, Pclass, Sex, Age, SibSp, Parch, Fare, Cabin, Embarked) %>% mutate(
  FirstClass = as.integer(Pclass == 1),
  SecondClass = as.integer(Pclass == 2),
  ThirdClass = as.integer(Pclass == 3),
  
  Male = as.integer(Sex == "male"),
  Female = as.integer(Sex == "female"),
  
  Cabins = case_when(
    is.na(Cabin) ~ 0,
    .default = unlist(str_split(Cabin, " ") %>% map(~ length(.x)))
  ),
  
  Cherbourg = as.integer(Embarked == "C"),
  Queenstown = as.integer(Embarked == "Q"),
  Southampton = as.integer(Embarked == "S"),
  
  Age = recode(Age, .missing = median(train$Age, na.rm=TRUE)),
  Fare = recode(Fare, .missing = median(train$Fare, na.rm=TRUE)),
  
  Survived = as.factor(Survived),
  
.keep="unused")
```

```{r}
#| echo: false

kable(head(titanic.training))
```

### Prepping Testing Data

```{r}

titanic.testing <- test %>% select(PassengerId, Pclass, Sex, Age, SibSp, Parch, Fare, Cabin, Embarked) %>% mutate(
  FirstClass = as.integer(Pclass == 1),
  SecondClass = as.integer(Pclass == 2),
  ThirdClass = as.integer(Pclass == 3),
  
  Male = as.integer(Sex == "male"),
  Female = as.integer(Sex == "female"),
  
  Cabins = case_when(
    is.na(Cabin) ~ 0,
    .default = unlist(str_split(Cabin, " ") %>% map(~ length(.x)))
  ),
  
  Cherbourg = as.integer(Embarked == "C"),
  Queenstown = as.integer(Embarked == "Q"),
  Southampton = as.integer(Embarked == "S"),
  
  Age = recode(Age, .missing = median(train$Age, na.rm=TRUE)),
  Fare = recode(Fare, .missing = median(train$Fare, na.rm=TRUE)),
  
  Survived = NA,
  
.keep="unused") %>% relocate(Survived, .after = PassengerId)
```

```{r}
#| echo: false

kable(head(titanic.testing))
```

## Models

### Random Forest

```{r}
#| output: false

rf <- randomForest(Survived~., data=titanic.training[,-1], na.action=na.roughfix)
rf.pred_training = rf %>% predict(titanic.training)
```

```{r}
#| echo: false

kable(table(titanic.training$Survived, rf.pred_training))
```

### Neural Net

```{r}
#| output: false

nn <- nnet(Survived~., data=titanic.training[,-1], size=13, na.action=na.roughfix)
nn.pred_training = nn %>% predict(titanic.training, type="class")
```

```{r}
#| echo: false

kable(table(titanic.training$Survived, nn.pred_training))
```

## Result

```{r}
titanic.testing$Survived = nn %>% predict(titanic.testing, type="class")
```

```{r}
#| echo: false

# titanic.testing %>% select(PassengerId, Survived) %>% write_csv("output.csv")

kable(head(titanic.testing))
```
